---
title: "Initial Korean covid 19"
author: "La Kayla Thompson"
date: "Summer 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```


### Introduction 

### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

$H_0$: There is no difference in COVID-19 mortality rates between males and females, the probability of dying is independent of sex.  

$H_A$: There is a difference in COVID-19 mortality rates between males and females, the probability of dying depends on sex.  


# Methods

Result is a categorical variable

Sex is also a categorical variable

result~sex, so we use the helper file cat~cat 


# Results

## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

``` 

The barchart shows that Korean males are more likely to die from covid then Korean females.

## Numerical results


## Inferential results 


```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical summary shows that more Korean females (98.7%) survived COVID-19 than Korean males (97%). This leads to a difference in death rates, with more Korean males (3%) dying from covid than Korean females (1.4%).

## Conclusions

